---
title: "p8105_hw1_nf2429"
author: "Nathalie Fadel"
date: "9/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#Problem 1
```{r}
set.seed(1)
Problem1_df = tibble(
  x = runif(10, 0, 5),
  vec_logical = (x > 2),
  vec_char = c("Carbon", "Hydrogen", "Nitrogen", "Helium", "Phosphorus", "Iodine", "Cesium", "Aluminum", "Oxygen", "Radium"),
  vec_factor = factor(c("Nonmetal", "Nonmetal", "Nonmetal", "Noble Gas", "Nonmetal", "Halogen", "Alkali Metal", "Post-transition metal", "Nonmetal", "Alkaline Earth Metal"))
) 
```

##Calculating Means
```{r}
mean(Problem1_df$x) 
mean(Problem1_df$vec_logical) 
mean(Problem1_df$vec_char) 
mean(Problem1_df$vec_factor) 
```
When we take the mean of variables x and vec_logical, we get actual numbers, since these variables are numeric. However, for vec_char and vec_factor, we get "NA"s for the outputs since the mean function does not work on character variables. 

##Numeric conversion
```{r}
as.numeric(Problem1_df$vec_logical) 
as.numeric(Problem1_df$vec_char) 
as.numeric(Problem1_df$vec_factor) 
```
This function on vec_logical outputs 1s and 0s to signify True and False. On the vec_character variable, we get "NA"s. On the vec_factor variable, the function assigned numbers 1-6 in alphabetical order to each of the categories.

##Other types of conversions
```{r}
as.factor(Problem1_df$vec_char)
as.numeric(as.factor(Problem1_df$vec_char))
as.character(Problem1_df$vec_factor)
as.numeric(as.character(Problem1_df$vec_factor))  
```
The output for the as.factor function on a character vector gives each of the 10 levels. When we put this function inside of the as.numeric function, it outputs numbers 1-10 that it has assigned to each of the levels in alphabetical order. The as.character function on the factor vector gives the categories for all 10 levels. When we try to convert from a character vector to numeric, the output gives "NA"s.

#Problem 2
```{r}
set.seed(1)
Problem2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  z = (x + y > 0),
z_num = as.numeric(z), #outputs 1s and 0s as True and False
z_fac = as.factor(z) #outputs the True and False as characters
)
```
The as.numeric function on the z logic vector outputs 1s and 0s to signify True and False, whereas the as.factor function on the z vector outputs True and False as characters.

This dataframe contains `r nrow(Problem2_df)` observations. The mean of x is `r mean(Problem2_df$x)`, and the median of x is `r median(Problem2_df$x)`. The proportion of TRUE cases is `r mean(Problem2_df$z)`. 

##Plots
```{r}
problem2_plot1 <- ggplot(Problem2_df, aes(x=x, y=y, colour=z)) + geom_point()  
ggsave("problem2_plot1.pdf", plot = problem2_plot1, height = 4, width = 6)   

problem2_plot1

ggplot(Problem2_df, aes(x=x, y=y, colour=z_num)) + geom_point() #blue color scale but the points are at one of two extremes of the spectrum.
ggplot(Problem2_df, aes(x=x, y=y, colour=z_fac)) + geom_point() #same as first plot
```


